---
title: "SDG 01: Poverty Projections"
output:
  xaringan::moon_reader:
    seal: false
    css: ["css/sfah.css", "default"]
    self_contained: false
    lib_dir: libs
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      titleSlideClass: ["left", "middle", "inverse"]
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(knitr)
library(scales)
library(ggpubr)
library(here)
library("hrbrthemes")

knitr::opts_chunk$set(
  warning    = FALSE,
  collapse   = TRUE,
  message    = FALSE,
  echo       = FALSE,
  fig.retina = 3,
  fig.width  = 10
)

# options(knitr.table.format = "html")
library(fontawesome) # from github: https://github.com/rstudio/fontawesome
xaringanExtra::use_xaringan_extra(c("tile_view", "animate_css", "tachyons"))
```

class: inverse, right, bottom
## SDG Atlas 2020

## Goal 01: Poverty Projections

---

class: middle, center, inverse

### Taking COVID into account
```{r distsim}
source(here("R", "SDG_01_covid_projections.R"))

pov_chart <-  function(x) {
  
  mpr <- x[, max(fgt, na.rm = TRUE)]
  mny <- x[, min(year, na.rm = TRUE)]
  mxy <- x[, max(year, na.rm = TRUE)]
  
  if (mxy >= 2019) {
    
    y <- x[year == 2019
    ][
      ,
      growth := "actual"
    ][,
      .(
        fgt   = mean(fgt, na.rm = TRUE),
        npoor = mean(npoor, na.rm = TRUE)
      ), 
      by = .(year, growth, scenario, povline)
    ]
    
    x <- rbindlist(list(x, y))
  }
  setorder(x, year, growth, scenario, povline)
  
  
  pg <- ggplot(x, 
         aes(
           x     = year, 
           y     = fgt, 
           color = growth
         )) +
    geom_line() +
    scale_y_continuous(
      labels = scales::percent,
      breaks = seq(0, mpr, round(mpr/10, digits = 3))
    ) +
    scale_x_continuous(
      # n.breaks = 8
      breaks = seq(mny, mxy, ceiling((mxy - mny + 1)/7))
    ) +
    labs(
      y     = "Poverty rate",
      x     = "",
      color = "scanario"
    ) +
    theme_minimal() +
    theme(
      legend.position = "none"
    )

  return(pg)
}
  

pl <- "1.9"
yr <- 2015
DC <- DC[povline == pl]

```
---
.left-column[
.f3[First we show the current poverty trend from 1993 to 2017, which is the time period used in this chapter]
]
.right-column[
```{r}
pov_chart(DC[scenario == "D"
             & year %between% c(1993, 2017) 
              ]
          )

```
]
---
.left-column[
.f3[Then, we zoom in to `r yr`-2021 and draw on pre-covid to 2021 to show what was the expected poverty rate before the pandemic]
]
.right-column[
```{r}
pov_chart(DC[scenario == "D"
             & year %between% c(yr, 2021) 
             & growth %chin% c("actual", "precovid")
              ]
          )


```
]
---
.left-column[
.f3[Then, we talk about the covid-baseline in the same period of time]
]
.right-column[
```{r}
pov_chart(DC[scenario == "D"
             & year %between% c(yr, 2021) 
             & growth %chin% c("actual", "precovid", "baseline")
              ]
          )


```
]
---
.left-column[
.f3[And finally we mention, with similar phrasing as in the blog, the covid-downside scenario]
]
.right-column[
```{r}
pov_chart(DC[scenario == "D"
             & year %between% c(yr, 2021) 
              ]
          )

```
]
---
.left-column[
.f3[Then, we mention that from 2022 onward it is expected that the global economy will start recovering so we extend all three series up to 2030]
]
.right-column[
```{r}

pov_chart(DC[scenario == "D"
             & year >= yr 
             ]
          )

```
]
---
.left-column[
.f3[Finally, we say that so far all the estimations are assuming neutral distribution growth, but when changes in the distribution are taken into account there is high uncertainty about what the poverty rate will be in 2030. Thus, we add a fan (top and bottom and shading all at once) based on worst downside and best baseline]
]
.right-column[
```{r}
DD <- 
  DC[
    year    >= (yr)
    ][
      ,
      `:=`(
        wr = fgt == max(fgt, na.rm = TRUE),
        bs = fgt == min(fgt, na.rm = TRUE)
      ), 
      by = .(year)
    ][
      wr == TRUE | bs == TRUE 
    ]

DD <- dcast(DD,
            year +  povline ~ scenario,
            value.var = c("fgt", "npoor"), 
            fun.aggregate = mean
    )



ggplot(data = DD,
       aes(
         x = year
       )) +
  geom_ribbon(aes(ymin = fgt_G,
                  ymax = fgt_B),
              fill  = "#69b3a2", 
              alpha = .7) +
  geom_line(aes(y = fgt_D), color = "black") +
  geom_line(aes(y = fgt_G), color = "black") +
  geom_line(aes(y = fgt_B), color = "black") +
  scale_y_continuous(
    labels = scales::percent,
    breaks = seq(0, max(DB$fgt_D, na.rm = TRUE), .05)
  ) +
  labs(
    y     = "Poverty rate",
    x     = "",
    color = "scanario"
  ) +
  theme_ipsum_rc() +
  theme(
    legend.position = "none"
  )
```
]
