---
title: "SDG 01: Poverty Projections"
output:
  xaringan::moon_reader:
    seal: false
    css: ["css/sfah.css", "css/fonts.css", "default"]
    self_contained: false
    lib_dir: libs
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      titleSlideClass: ["left", "middle", "inverse"]
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(knitr)
library(scales)
library(ggpubr)
library(here)
library("hrbrthemes")

knitr::opts_chunk$set(
  warning    = FALSE,
  collapse   = TRUE,
  message    = FALSE,
  echo       = FALSE,
  fig.retina = 3,
  fig.width  = 10
)
# knitr::opts_knit$set(
#   root.dir = ".."
# )
# options(knitr.table.format = "html")
library(fontawesome) # from github: https://github.com/rstudio/fontawesome
xaringanExtra::use_xaringan_extra(c("tile_view", "animate_css", "tachyons"))
```

class: inverse, right, bottom
## SDG Atlas 2020

## Goal 01: Poverty Projections

???
Intcomments

---

class: middle, center, inverse

### Taking COVID into account
```{r distsim}
source(here("R", "SDG_01_covid_projections.R"))

pov_chart <-  function(x) {
  
  mpr <- x[, max(fgt, na.rm = TRUE)]
  mny <- x[, min(year, na.rm = TRUE)]
  mxy <- x[, max(year, na.rm = TRUE)]
  
  if (mxy >= 2019) {
    
    y <- x[year == 2019
    ][
      ,
      growth := "actual"
    ][,
      .(
        fgt   = mean(fgt, na.rm = TRUE),
        npoor = mean(npoor, na.rm = TRUE)
      ), 
      by = .(year, growth, scenario, povline)
    ]
    
    x <- rbindlist(list(x, y))
  }
  setorder(x, year, growth, scenario, povline)
  
  
  pg <- ggplot(x, 
         aes(
           x     = year, 
           y     = fgt, 
           color = growth
         )) +
    geom_line() +
    scale_y_continuous(
      labels = scales::percent,
      breaks = seq(0, mpr, round(mpr/10, digits = 3))
    ) +
    scale_x_continuous(
      # n.breaks = 8
      breaks = seq(mny, mxy, ceiling((mxy - mny + 1)/7))
    ) +
    labs(
      y     = "Poverty rate",
      x     = "",
      color = "scanario"
    ) +
    theme_minimal() +
    theme(
      legend.position = "none"
    )

  return(pg)
}
  

pl <- "1.9"
yr <- 2015
DC <- DC[povline == pl]

```
---
.left-column[
Past data 1993 to 2017, axis to 1993-2017
]
.right-column[
```{r}
pov_chart(DC[scenario == "D"
             & year %between% c(1993, 2017) 
              ]
          )

```
]
---
.left-column[
Zoom in to `r yr`-2021, draw on pre-covid to 2021
]
.right-column[
```{r}
pov_chart(DC[scenario == "D"
             & year %between% c(yr, 2021) 
             & growth %chin% c("actual", "precovid")
              ]
          )


```
]
---
.left-column[
Draw on covid-baseline to 2021
]
.right-column[
```{r}
pov_chart(DC[scenario == "D"
             & year %between% c(yr, 2021) 
             & growth %chin% c("actual", "precovid", "baseline")
              ]
          )


```
]
---
.left-column[
Draw on covid-downside to 2021
]
.right-column[
```{r}
pov_chart(DC[scenario == "D"
             & year %between% c(yr, 2021) 
              ]
          )

```
]
---
.left-column[
Extend all axis and all three series `r yr`-2030
]
.right-column[
```{r}

pov_chart(DC[scenario == "D"
             & year >= yr 
             ]
          )

```
]
---
.left-column[
Add fan (top and bottom and shading all at once) based on worst-downside and best-covid-baseline
]
.right-column[
```{r}
DD <- 
  DC[
    year    >= (yr)
    ][
      ,
      `:=`(
        wr = fgt == max(fgt, na.rm = TRUE),
        bs = fgt == min(fgt, na.rm = TRUE)
      ), 
      by = .(year)
    ][
      wr == TRUE | bs == TRUE 
    ]

DD <- dcast(DD,
            year +  povline ~ scenario,
            value.var = c("fgt", "npoor"), 
            fun.aggregate = mean
    )



ggplot(data = DD,
       aes(
         x = year
       )) +
  geom_ribbon(aes(ymin = fgt_G,
                  ymax = fgt_B),
              fill = "#69b3a2") +
  geom_line(aes(y = fgt_D), color = "black") +
  geom_line(aes(y = fgt_G), color = "black") +
  geom_line(aes(y = fgt_B), color = "black") +
  scale_y_continuous(
    labels = scales::percent,
    breaks = seq(0, max(DB$fgt_D, na.rm = TRUE), .05)
  ) +
  labs(
    y     = "Poverty rate",
    x     = "",
    color = "scanario"
  ) +
  theme_ipsum_rc() +
  theme(
    legend.position = "none"
  )
```
]
