---
title: "Share of population below the 50% of the median"
output:
  xaringan::moon_reader:
    seal: false
    css: ["css/sfah.css", "css/fonts.css", "default"]
    self_contained: false
    lib_dir: libs
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%"
      ratio: "16:9"
      titleSlideClass: ["left", "middle", "inverse"]
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library("knitr")
library("scales")
library("xaringanthemer")
library("here")
library("pins")
library("tidyverse")
library("plotly")
library("data.table")


knitr::opts_chunk$set(
  warning = FALSE,
  collapse = TRUE,
  message = FALSE,
  echo = FALSE,
  fig.retina = 3,
  fig.width = 10
)
knitr::opts_knit$set(
  root.dir = ".."
)
# options(knitr.table.format = "html")
library(fontawesome) # from github: https://github.com/rstudio/fontawesome
xaringanExtra::use_xaringan_extra(c("tile_view", "animate_css", "tachyons"))

```

class: inverse, right, bottom
## Share of population below 50% of Median

## Analysis by groups of countries

???
Initial comments
```{r initcal}
if (length(ls(pattern = "ld_50pm_1")) == 0 ) { # load df?
  source(here("R", "SDG_10_50percent_median.R"))
  ld_50pm_1 <-  1
}
```

---
.pull-left[
.f1.pv6.orange[
Target 10.2
]
]
.pull-right[
Target 10.2 seeks to .red[*empower and promote the social, economic and political inclusion of all, irrespective of age, sex, disability, race, ethnicity, origin, religion or economic or other status.*] One way to measure if this target is being met is by reducing the .purple[*Proportion of people living below 50 percent of median income.*] This is sometimes called relative poverty, but it is indeed more related to a measure of inequality and social inclusion than to a measure of deprivation. 

People who fall below this cutoff are considered relatively deprived in the sense that their socioeconomic status is substantially lower than the one of the better off in a society. Irrespective from the general economic development of a society, if the share of the population that falls below 50% of the median does not change over time, we can this society lack .purple[**social, economic and political inclusion.**]
]
---
.left-column[
By using comparable data, we found those countries for which there is data available after `r first_year` for two points that are at least `r min_diff` years apart. 

In this chart, the countries at the top are those that reduce their social deprivation the most, whereas the ones at the bottom increase the share of people living below 50% of the median
]
.right-column[
```{r}
hc_50med(dw)
```
]

---
.left-column[
.orange.f2[Low-income]

And this is regular text
]
.right-column[
```{r}
hc_50med(dw, incomegroup == "Low income")
```
]

---
.left-column[
.orange.f2[Low-middle-income]
]
.right-column[
```{r}
hc_50med(dw, incomegroup == "Lower middle income")
```
]

---
.left-column[
.red.f2[Upper-middle-income]
]
.right-column[
```{r}
hc_50med(dw, incomegroup == "Upper middle income")
```
]

---
.left-column[
.red.f2[High-income]
]
.right-column[
```{r}
hc_50med(dw, incomegroup == "High income")
```
]

---
.left-column[
.blue.f2[EAP]
]
.right-column[
```{r}
hc_50med(dw, region == "EAP")
```
]

---
.left-column[
.green.f2[ECA]
]
.right-column[
```{r}
hc_50med(dw, region == "ECA")
```
]

---
.left-column[
.purple.f2[LAC]
]
.right-column[
```{r}
hc_50med(dw, region == "LAC")
```
]

---
.left-column[
.pink.f2[EAP]
]
.right-column[
```{r}
hc_50med(dw, region == "MNA")
```
]

---
.left-column[
.navy.f2[EAP]
]
.right-column[
```{r}
hc_50med(dw, region == "SAS")
```
]

---
.left-column[
.gold.f2[EAP]
]
.right-column[
```{r}
hc_50med(dw, region == "SSA")
```
]

---
.left-column[
.dark-red.f2[EAP]
]
.right-column[
```{r}
hc_50med(dw, region == "OHI")
```
]


