---
title: "SDG 10"
output:
  xaringan::moon_reader:
    seal: false
    css: ["css/sfah.css", "css/fonts.css", "default"]
    self_contained: false
    lib_dir: libs
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      titleSlideClass: ["left", "middle", "inverse"]
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(knitr)

knitr::opts_chunk$set(
  warning = FALSE,
  collapse = TRUE,
  message = FALSE,
  echo = FALSE,
  fig.retina = 3,
  fig.width = 10
)
knitr::opts_knit$set(
  root.dir = ".."
)
# options(knitr.table.format = "html")
library(fontawesome) # from github: https://github.com/rstudio/fontawesome
xaringanExtra::use_xaringan_extra(c("tile_view", "animate_css", "tachyons"))
```

class: inverse, right, bottom
## SDG Atlas 2020

## Goal 10: Reduce inequality within and among countries

???
Intcomments

---

class: middle, center, inverse

### Story of inequality
```{r distsim}
source("R/utils.R")
source("R/SDG_10_lorenz.R")
```
---

.left-column[
### Start with perfect equality

.big[Everybody has the same amount of money]
]
.right-column[
```{r pe1}

set.seed(1010111)
plt <- sample(palette, size = 4)

breaks <- df %>%
  group_by(src) %>%
  summarise(
    gini = gini(y,w)
  ) %>%
  arrange(-gini) %>%
  pull(src)

f <- df %>% 
  group_by(src) %>%  
  filter(
      abs(cw-.25) == min(abs(cw-.25))
    | abs(cw-.5)  == min(abs(cw-.5))
    | abs(cw-.75) == min(abs(cw-.75))
      ) %>% 
  select(src, x2 = cw, y1 = cy) %>% 
  mutate(y2 = y1,
         x1  = 0) 


pe <- ggplot(data=filter(df, src %in% breaks[4]),
             aes(x=cw,
                 y=cy,
                 color = src)) +
  geom_line() +
  scale_x_continuous(name="Cumulative share of population", limits=c(0,1)) +
  scale_y_continuous(name="Cumulative share of welfare", limits=c(0,1)) +
  theme_classic() +
  scale_color_manual(values= plt,
                     breaks = breaks)
pe

```
]

---
.left-column[
### Start with perfect equality

.big[welfare and population cummulate at the same pace. So, the bottom 25 percent of the population has 25% of the total income]
]
.right-column[
```{r pe2}

sgm <- data.frame(x1 = 0, x2 = 0.25, y1 = 0.25, y2 = .25)

pe2 <- pe +
  geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2), colour = palette[10], data = sgm) +
  geom_segment(aes(x = y1, y = x1, xend = y2, yend = x2), colour = palette[10], data = sgm)
pe2
```
]

---
.left-column[
### Start with perfect equality

.big[welfare and population cummulate at the same pace. So, the bottom 25 percent of the population has 25% of the total income. The bottom 50 of the population has 50 percent of total income, and so on... 
]

]
.right-column[
```{r pe3}

sgm2 <- data.frame(x1 = 0, x2 = .5, y1 = .5, y2 = .5)

pe3 <- pe2 +
  geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2), colour = palette[10], data = sgm2) +
  geom_segment(aes(x = y1, y = x1, xend = y2, yend = x2), colour = palette[10], data = sgm2)

pe3
```
]

---
.left-column[
### A very unequal country


]
.right-column[
```{r zaf1}
pzaf <- ggplot(data= filter(df, src %in% breaks[c(1, 4)]),
       aes(x=cw,
           y=cy,
           color = src)) +
  geom_line() +
  scale_x_continuous(name="Cumulative share of population", limits=c(0,1)) +
  scale_y_continuous(name="Cumulative share of welfare", limits=c(0,1)) +
  theme_classic() +
  scale_color_manual(values= plt, 
                     breaks = breaks)
pzaf
```
]
---
.left-column[
### A very unequal country


]
.right-column[
```{r zaf2}

f1 <- f %>% 
  filter(src  %in% breaks[1]) %>% 
  slice(1)


pzaf1 <- pzaf +
  geom_segment(aes(x    = x1, 
                   y    = y1, 
                   xend = x2, 
                   yend = y2
                   ), 
               colour = palette[10], 
               data = f1
               ) +
  geom_segment(aes(x    = x2, 
                   y    = x1, 
                   xend = x2, 
                   yend = y2
                   ), 
               colour = palette[10], 
               data = f1
               )
pzaf1
```
]
---
.left-column[
### A very unequal country


]
.right-column[
```{r zaf3}

f2 <- f %>% 
  filter(src  %in% breaks[1]) %>% 
  slice(2)

pzaf2 <- pzaf1 +
  geom_segment(aes(x    = x1, 
                   y    = y1, 
                   xend = x2, 
                   yend = y2
                   ), 
               colour = palette[10], 
               data = f2
               ) +
  geom_segment(aes(x    = x2, 
                   y    = x1, 
                   xend = x2, 
                   yend = y2
                   ), 
               colour = palette[10], 
               data = f2
               )
pzaf2
```
]
---
.left-column[
### A very unequal country


]
.right-column[
```{r zaf4}

f3 <- f %>% 
  filter(src  %in% breaks[1]) %>% 
  slice(3)

pzaf3 <- pzaf2 +
  geom_segment(aes(x    = x1, 
                   y    = y1, 
                   xend = x2, 
                   yend = y2
                   ), 
               colour = palette[10], 
               data = f3
               ) +
  geom_segment(aes(x    = x2, 
                   y    = x1, 
                   xend = x2, 
                   yend = y2
                   ), 
               colour = palette[10], 
               data = f3
               )
pzaf3
```
]
---
.left-column[
### What if the T10 has everything?

]
.right-column[
```{r BT}
pfin <- ggplot(data= filter(df, src %in% breaks[c(1, 3, 4)]),
       aes(x=cw,
           y=cy,
           color = src)) +
  geom_line() +
  scale_x_continuous(name="Cumulative share of population", limits=c(0,1)) +
  scale_y_continuous(name="Cumulative share of welfare", limits=c(0,1)) +
  theme_classic() +
  scale_color_manual(values= plt, 
                     breaks = breaks)

```
]
