---
title: "SDG 10"
output:
  xaringan::moon_reader:
    seal: false
    css: ["css/sfah.css", "css/fonts.css", "default"]
    self_contained: false
    lib_dir: libs
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      titleSlideClass: ["left", "middle", "inverse"]
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(knitr)
library(data.tree)

knitr::opts_chunk$set(
  warning = FALSE,
  collapse = TRUE,
  message = FALSE,
  echo = FALSE,
  fig.retina = 3,
  fig.width = 10
)
knitr::opts_knit$set(
  root.dir = ".."
)
# options(knitr.table.format = "html")
library(fontawesome) # from github: https://github.com/rstudio/fontawesome

```

class: inverse, right, bottom
## SDG Atlas 2020

## Goal 10: Reduce inequality within and among countries

---
```{r ginichange}
source("R/SDG_10_Gini_Change.R")

```

class: center, middle

## .saltinline[What can we say about Global inequality during the last 15 or 20 years?]

--

### .saltinline[Has it gone up or down?]

---
class: center, middle

## .fatinline[Not an easy answer]

--

### .fatinline[In contrast to poverty, inequality measures, Gini in particular, are not additive decomposable. That is, the average of Ginis is not equal to the total Gini.]

---

## .acidinline[Total Gini]

.big[If I pull the welfare aggregates of all the countries and calculate a global Gini as if it were a single country, what does that tell me?]
--

.big[It does not say much because whether it goes up or down over time, I have no way to explain that movement:]

--
.big[
- The distributional changes across countries may cancel each other 
- The welfare between countries is not comparable (e.g., cons and income)
- I can't see where the action is happening. 
- **There is no data available to create a representative set of countries in any year to build a comparable trend of global inequality over time**]
---
## .acidinline[Average of Ginis]

--
.big[This should be simple, it is just estimating how inequality, in average, has moved over time]

--

.big[It is not that simple because, 
- Again, we don't have data for all the countries in all the years, so any two sets of countries in two years will be different to each other. 
- Lackner et. al (2019) found no more than 43 countries with which we could calculate an average of the Ginis over time, but that set is not representative of the whole world.
]
---
class: center, middle
## .saltinline[So, If we want a simple measure of global inequality, how can we overcome the problem of lack of data?]

--

.big[.heat[We could interpolate ] `r emo::ji("+1")`]

---
.left-column[
.big[.acidinline[Without interpolation]]

By using **only** available data, the *average* Gini coefficient decreased steadly from the mid 90's to 2010. Since then, the trend reverted and the average inequality is increasing rapidly.
]

.right-column[
```{r giniav-noint}
p_av_g +
  geom_vline(xintercept = c(1996, 2010, 2013),
             color = "gray70")
```
]

---
.left-column[
.big[.acidinline[With interpolation]]

However, when we interpolate linearly to account for countries with missing data in some years but with at least two data points five years appart, the story changes. 
]

.right-column[
```{r giniav}
p_av_g_int +
  geom_vline(xintercept = c(1996, 2010, 2013),
             color = "gray70")
```
]

---

class: center, middle
## .saltinline[Ok, if I can't say anything about global inequality in general, can I say something about the countries?]

--

### .saltinline[can we know which countries are the most unequal?]

--

### .saltinline[can we know how has inequality evolved in each country?]

---
class: center, middle

## .heatinline[Of course! we can answer that]

---

## .heatinline[but...]

--

.big[
- It is not exciting
- It has already been done
- It's about it. There's no much we can say
]

--

.center[.heat[Let's see it]]
---

.left-column[
.big[.acidinline[Gini coef. by country in circa 2000-2018]]

Here we have all the countries of the world and their corresponding Gini coefficient of the closest year to 2000 and 2018. 
]
.right-column[
```{r gini45}
p_g
```
]

---

.left-column[
.big[.acidinline[Gini coef. by country in circa 2000-2018 (LAC)]]

LAC is the region that consistently has reduced inequality in most countries. This makes sense because inequality is measured with income-based welfare, which is by definition higher than consumption-based welfare,  so there is more room to improve.
]
.right-column[
```{r gini45lac}
p_g +
  ggforce::geom_mark_hull(
    aes(filter = region == 'LAC',
        fill = region,
        label = region),
    show.legend = FALSE,
    expand = unit(2, "mm"),
    concavity = 1
  )
``` 
]

---
.left-column[
.big[.acidinline[Gini coef. by country in circa 2000-2018 (SSA)]]

When we look at other regions, we don't find the same patterns.
]
.right-column[
```{r gini45SSA}
p_g +
  ggforce::geom_mark_hull(
    aes(filter = region == 'SSA',
        fill = region,
        label = regionname),
    show.legend = FALSE,
    expand = unit(2, "mm"),
    concavity = 1
  )
```
]

---
.left-column[
.big[.acidinline[Gini coef. by country in circa 2000-2018 (EAP)]]

When we look at other regions, we don't find the same patterns.
]
.right-column[
```{r gini45EAP}

p_g +
  ggforce::geom_mark_hull(
    aes(filter = region == 'EAP',
        fill = region,
        label = regionname),
    show.legend = FALSE,
    expand = unit(2, "mm"),
    concavity = 1
  )
```
]
---

class: center, middle

## .heatinline[We could also highlight by income groups instead]

---
.left-column[
.big[.acidinline[Gini coef. by country in circa 2000-2018 (Low income)]]
]
.right-column[
```{r gli}
p_g_i +
  ggforce::geom_mark_hull(
    aes(filter = incomegroup == 'Low income',
        fill = incomegroup,
        label = incomegroup),
    show.legend = FALSE,
    expand = unit(2.5, "mm"),
    concavity = 1)

```
]

---

.left-column[
.big[.acidinline[Gini coef. by country in circa 2000-2018 (Lower middle income)]]
]
.right-column[
```{r glmi}
p_g_i +
  ggforce::geom_mark_hull(
    aes(filter = incomegroup == 'Lower middle income',
        fill = incomegroup,
        label = incomegroup),
    show.legend = FALSE,
    expand = unit(2.5, "mm"),
    concavity = 1)

```
]

---

.left-column[
.big[.acidinline[Gini coef. by country in circa 2000-2018 (Upper middle income)]]
]
.right-column[
```{r gumi}
p_g_i +
  ggforce::geom_mark_hull(
    aes(filter = incomegroup == 'Upper middle income',
        fill = incomegroup,
        label = incomegroup),
    show.legend = FALSE,
    expand = unit(2.5, "mm"),
    concavity = 1)

```
]

---

.left-column[
.big[.acidinline[Gini coef. by country in circa 2000-2018 (High income)]]
]
.right-column[
```{r ghi}

p_g_i +
  ggforce::geom_mark_hull(
    aes(filter = incomegroup == 'High income',
        fill = incomegroup,
        label = incomegroup),
    show.legend = FALSE,
    expand = unit(2.5, "mm"),
    concavity = 2)

```
]

---
### .saltinline[Ok, inequality is decreasing in most of LAC countries and goes all over the place for the rest of countries. So what?...] 

--

- .big[What does a reduction in the global Gini actually says?]

--

- .big[Where is the action happening within the distribution of the countries?]

--

- .big[From a technical point of view, why poor countries have higher inequality than rich countries?]

--

- .big[Can we say that one country is more unequal than another by just comparing thier Ginis?]

--

- .big[How similar are the welfare levels of two countries with similar inequality?]


---
### .heatinline[Since it is difficult to come up with a compelling story for global inequality, we could do two things]

--

1. .big[We could shed some light on the following questions]

--

2. .big[We could provide the user with a unique experience. ]
--
.big[A set of tools to help them understand their position in the welfare distribution of the world]

---
class: inverse, center, middle
## Another way to see inequality

```{r p1090-calc}
source("R/SDG_10_p10p90p50.R")
```
---

.left-column[
Here, countries are organized according to their median, from lowest to highest.
]
.right-column[
```{r p50}
p_p50
```
]

---
```{r medianratio}
h50l50 <- dfc_p50[["p50"]][nrow(dfc_p50)]/dfc_p50[["p50"]][1]
h50l50 <- floor(h50l50)
```
.left-column[
Here, countries are organized according to their median, from lowest to highest.

From `r dfc_p50[["countryname"]][1]` to `r dfc_p50[["countryname"]][nrow(dfc_p50)]`. 

By just looking at the chart we can see that the *typical person* in `r dfc_p50[["countryname"]][nrow(dfc_p50)]` is `r h50l50` times richer than the *typical person* in `r dfc_p50[["countryname"]][1]`

]
.right-column[
```{r p50l-lh}
p_p50 +
  ggrepel::geom_label_repel(
    data = dfc_p50lh,
    aes(label = text,
        fill  = region),
    show.legend = FALSE,
    force = 20,
    box.padding = 1.2,
    # max.overlaps = 2,
    segment.curvature = 0.5,
    # segment.ncp = 3,
    # segment.angle = 20,
    nudge_y = 5
  )

```
]
---
.left-column[
Here, countries are organized according to their median, from lowest to highest.

From `r dfc_p50[["countryname"]][1]` to `r dfc_p50[["countryname"]][nrow(dfc_p50)]`. 

By just looking at the chart we can see that the *typical person* in `r dfc_p50[["countryname"]][nrow(dfc_p50)]` is `r h50l50` times richer than the *typical person* in `r dfc_p50[["countryname"]][1]`

]
.right-column[
```{r p50-all}
p_p50 +
  ggrepel::geom_label_repel(
    data = dfc_p50,
    aes(label = text,
        fill  = region),
    show.legend = FALSE,
    force = 20,
    box.padding = 1.2,
    # max.overlaps = 2,
    segment.curvature = 0.5,
    # segment.ncp = 3,
    # segment.angle = 20,
    nudge_y = 5
  )

```
]
---
.left-column[
Now, let's remove the labels and add the monetary distance between the 10th and 90th percentiles. 
]
.right-column[
```{r p50l2}
p_p50
```
]
---
.left-column[
Now, let's remove the labels and add the monetary distance between the 10th and 90th percentiles. 

]
.right-column[
```{r p1090g}
p_p10p90
```
]

---
.left-column[
As in this random case, the upper limit of the bar represents the 90th percentile of the distribution, whereas the lower limit represents the 10th percentile of the distribution. 
]
.right-column[
```{r p1090gc}
p_p10p90lc
```
]
---
.left-column[
By looking at the monetary distance between p90 and p10 of these two countries, 

.heatinline[which of them is more unequal? A or B?]
]
.right-column[
```{r 2cx}
p_p10p90_2cx
```
]
---
.left-column[
By looking at the monetary distance between p90 and p10 of these two countries, 

.heatinline[which of them is more unequal? A or B?]

.heatinline[surprised?]
]
.right-column[
```{r 2cx2}
p_p10p90_2cx + geom_text(aes(label = paste("Gini:\n", round(gini, digits = 3))),
                         nudge_x = -.1,
                         nudge_y = 2.8)
```
]
---
.left-column[
Even though the length of the bars makes it look like there is more inequality within rich countries than within poor ones, it is just a visual illusion.
]
.right-column[
```{r p1090g2}
p_p10p90
```
]

---
```{r}
sp90   <- paste0("$",round(dfc_rep[["p90"]], digits = 2))
sp10   <- paste0("$",round(dfc_rep[["p10"]], digits = 2))
sr9010 <- round(dfc_rep[["r9010"]], digits = 2)

```


.left-column[
Even though the length of the bars makes it look like there is more inequality within rich countries than within poor ones, it is just a visual illusion.

Take for instance `r dfc_rep[["countryname"]][1]`, whose 90th percentile (`r sp90[1]`) is `r sr9010[1]` times higher than its 10th percentile (`r sp10[1]`). 
]
.right-column[
```{r p1090gl2}
p_p10p902
```
]

---

.left-column[
Even though the length of the bars makes it look like there is more inequality within rich countries than within poor ones, it is just a visual illusion.

Take for instance `r dfc_rep[["countryname"]][1]`, whose 90th percentile (`r sp90[1]`) is `r sr9010[1]` times higher than its 10th percentile (`r sp10[1]`). 

In contrast, `r dfc_rep[["countryname"]][nrow(dfc_rep)]`'s  90th percentile (`r sp90[nrow(dfc_rep)]`) is only `r sr9010[nrow(dfc_rep)]` times higher than its 10th percentile (`r sp10[nrow(dfc_rep)]`). 
]
.right-column[
```{r p1090gl3}
p_p10p903
```
]

---
.left-column[
Yet, this chart does not imply that the richer the country the lower its inequality levels.  

]
.right-column[
```{r p1090-4}
p_p10p903
```
]
---
.left-column[
Yet, this chart does not imply that the richer the country the lower its inequality levels. 

If we keep the countries sorted by their median and plot the p90/p10 ration in the Y axis, the current sorting is completely lost.
]
.right-column[
```{r rj}
p_r9010_j
```
]
---
.left-column[
If we reorganize the countries according to their p90/p10 ratio, we see a strong re-ranking. So, you  may wondering

- is the Gini highly correlated to the p90/p10 ratio?

- is there any relationship between welfare and inequality? 
]
.right-column[
```{r ro}
p_r9010_o
```
]
---
.left-column[
.heatinline[Gini vs P90/P10]

Even though there is a sort of quadratic relation between the Gini and the 90/10 ratio, it is difficult to explain the high variability. 
]
.right-column[
```{r p9010g}
p_9010_g
```
]
---
.left-column[
.heatinline[Gini vs Median]

The richer countries are not necessarily the less unequal. 
]
.right-column[
```{r p9010med}
p_med_9010
```
]
---

class: center, middle, inverse

## What is happening at the tails?

---

class: center, middle

## What happen to inequality ranking if we removed the top 5% of all the countries?

```{r conditionalgini}
source("R/SDG_10_conditional_ginis.R")
```

---
```{r g0095-1}
p_0095_1
```

---

```{r g0095-2}
p_0095_2
```

---
.left-column[
The most remarkable change in ranking happens in `r names(cc1)[1]`, followed by `r names(cc1)[2]`
]
.right-column[
```{r g0095-3}
p_0095_3
```
]

---
.left-column[
The most remarkable change in ranking happens in `r names(cc1)[1]`, followed by `r names(cc1)[2]` and `r add_and(names(cc1)[3:5])`
]
.right-column[
```{r g0095-4}
p_0095_4
```
]

---
.left-column[

Even though the Gini coefficient is lower for all the countries after removing the top 5%, all high income countries (with the exception of *Seychelles*) descended in the global inequality ranking.
]
.right-column[
```{r p_0095-h}
p_0095_h
```
]

---
.left-column[

... and the same happens to upper-middle-income countries
]
.right-column[
```{r p_0095-u}
p_0095_um
```
]


---
.left-column[

Yet, if some countries move down in the global ranking, others must go up. This the case for low-income countries
]
.right-column[
```{r p_0095-l}
p_0095_l
```
]


---
.left-column[

Yet, if some countries move down in the global ranking, others must go up. This the case for low-income countries and lower-middle-income countries
]
.right-column[
```{r p_0095-lm}
p_0095_lm
```
]
---

class: middle, center

### the top of the distribution highly affects the ranking of global inquality

--

### .saltinline[So, how can we know if a particular country is more unequal than other? By using the Gini?] 

--

### .heatinline[Not really, as it might be reflecting just the effect of the top, hidding the rest of the distribution.]

--

### The main problem is not the inequality wihtin countries but between countries.

---
class: middle, center

.big[let's take the median of each country, which is not affected by the outliers of the top, and plot a Gini coefficient for the world over time.]


---

```{r ginimed}
source("R/SDG_10_global_gini.R")
```

.left-column[
Under this perspective, where we take a measure of central tendency that is not affected by outliers (say, within-country inequality), we can see that between-country in the world has increased
]
.right-column[
```{r ginimed-u}
p_gini_med
```
]

---

class: middle, center

.big[Ok, so that is the evolution of inequality using the median. What about the rest of the distribution?]

---


