---
title: Goal 1. End poverty in all its forms everywhere
subtitle: Your subtitle
author: R.Andres Castaneda
date: Created `r Sys.Date()`
output:
  redoc::redoc:
    reference_docx: "template.docx"
    highlight_outputs: TRUE
    margins: 1 
    line_numbers: FALSE 
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  warning = FALSE,
  collapse = TRUE,
  message = FALSE,
  echo = FALSE,
  fig.retina = 3
)
knitr::opts_knit$set(
  root.dir = ".."
)
```

```{r calculations}

source("R/SDG_01_data.R")
source("R/SDG_01_charts.R")

# No. of poor
npoor <- wld %>% 
  arrange(year) %>% 
  filter(year  %in% c(1993, 2015)) %>% 
  select(poor_pop) %>% 
  round(digits = 0) %>% 
  pull()

np <- (npoor[1] - npoor[2])


# Headcounts
hc <- wld %>% 
  arrange(year) %>% 
  filter(year  %in% c(1993, 2015)) %>% 
  select(headcount) %>% 
  {round(.*100, digits = 1)} %>% 
  pull()

tmp = "TEMP"
x <- "x"

#--------- regional poveryt

r <- povcalnet_wb() 


#--------- CHN and IND poor population
np_ic <- cty2 %>% 
  filter(year  %in% c(1993, 2015)) %>% 
  group_by(year) %>% 
  summarise(pop = sum(poor_pop)) %>% 
  ungroup() %>% 
  arrange(-year) %>% 
  summarise(diff = diff(pop)) 

sh_np <- prettyNum((np_ic/10)/np , digits = 3)

```

# The story behind Global poverty

The first target of Goal 1 of the SDG is to eliminate extreme poverty by 2030.
Based on the historical trend of the most recent data, the World seems to be
going in the right direction to meet the goal. In 1990, `r hc[1]`
percent of the population (`r prettyNum(npoor[1]/1000, digits = 2)` billion people) lived with
less than \$1.9 a day (PPP 2011 prices). 25 years later,
`r prettyNum((np/npoor[1])*100, digits = 2)` percent of theme escaped extreme poverty, leaving
behind `r prettyNum(npoor[2], digits = 3)` million people (`r hc[2]` of the
population).

If we assume that the trend of poverty reduction will remain the same over the
next decade as it has been during the last 25 years, Goal 1 would be met by
**2024!**. Unfortunately, this optimistic scenario is not support by the data.
Let's see why.

First, the underlying data of the Global poverty rate in a given year are the
poverty rates of all the countries in the world. In some of them, the poverty
rate is high, but in others it is low. The dispersion of poverty rates in a
given years is so wide that scale at which global poverty is usually analized
must be increased to make room for all the poverty rates to fit the chart. At
this scale, the decreasing pace of global poverty does not seem that rapid
anymore.

How could it be then that the dispersion of poverty rates has remained
relatively constant over time and yet global poverty rate has decreased? This is
so because global poverty is a population-weigthed average of country poverty
rates. The larger the country the more it weighs in the global average.

By looking closely at the underlying data, we find that the world-wide poverty
story of the last 25 five years is mainly explained by the economic performance
of two countries: China and India. Out of the `r prettyNum(np, digits = 2)` billion
people that escaped poverty from 1990 to 2015, `r np_ic` million
(`r sh_np` percent) is due to these two countries. The other
`r 100- as.numeric(sh_np)` percent is the contribution of the rest of the world.

```{r china}
chn_mm <- cty2 %>% 
    group_by(countrycode) %>% 
    slice(which.min(year), which.max(year)) %>% 
    filter(countrycode == "CHN") %>% 
    select(headcount) %>% 
    pull()
```

China, in particular, has reduced its poverty rate from
`r prettyNum(chn_mm[[1]]*100, digits = 2)` percent in 1993 to `r prettyNum(chn_mm[[2]]*100, digits = 2)` percent
in 2015. The number of extreme poor in China is so small today that it bearly
counts in the global average. India, on the other, though it has reduced by half
its poverty rate, it still has some room to contribute to reduce global poverty.

During the same period, other countries have reduced their poverty rate as well,
but many small countries still have poverty rates that are way higher than the
world average. Unfortunately, most of those countries belong to Subsaharan
Africa and in many of them poverty has increased during during the last decade.

Other regions like Latin America and Europe and Central Asia are not doing that
bad. Most of the their countries have poverty rates below the global average and
the great majority has reduced their povety rates over the same period.

Summing up, this is how the world looks today. The higest poverty rates are
concentrated in Africa and South Asia and only two countries, India and Nigeria.
have `r x` of the poor. Yet, How do we expect it will look
like inthe future?

According to `r x`, if we assume an accelerated economic
growth and a constant reduction of inequality in all the countries, the best
projection of global poverty still fails to eliminate global poverty by 2030. In
contrast, projections of poverty rates that assume poor economic growth and
rising welfare inequality are about `r scales::percent(max(pty$headcount), suffix = "")` percent.

# Different trends across countries

If we assume that each country were to keep the same poverty-reduction trend of the last 20 years, we could divide the countries in the world in three groups. Those that will eventually eliminate poverty, those that won't eliminate poverty, and those that we jsut don't know. 

Countries whose poverty trends have been negative during the last years will eventually eliminate poverty as long as the trend remains unchanged. The current poverty-reduction trends are not predictor of poverty levels in the future, as they do not take into account poitical, economic, or exogenous events that affect poverty levels. However, the trend provides us with sense of the current pace at which each country is reducing poverty. This pace may change in the future, but if it were to continue indefinetly as is, countries that are currently reducing poverty levels will eventually eliminate poverty completely. The question is the, when would that happen?

```{r graduation}
lower_year <- 2020

cty_g <- cty_p %>% 
  filter(year > lower_year) %>% 
  arrange(-year) 

shr_grd <- tabyl(cty_g, region)  # graduation share

grd_per <- shr_grd %>%           # Africa share
  mutate(
    percent = scales::percent(percent, 
                              suffix = "", 
                              accuracy = 0.1)
    ) %>% 
  select(-n)

avg_year <- cty_g %>%        # average year
  group_by(region) %>% 
  summarise(ymean = floor(mean(year)))

```


Among all the countries that eventually would eliminate poverty after `r lower_year`, `r grd_per[grd_per$region == "SSA", "percent"]` percent are from Sub Saharan Africa, `r grd_per[grd_per$region == "LAC", "percent"]` percent are from Latin America, and `r grd_per[grd_per$region == "EAP", "percent"]` percent are from East Asia and Pacific. This proportion is expected as the number of countries in Africa is way larger than in any other region, but what is unexpected is the their year of graduation. In average, African countries would graduate in `r avg_year[avg_year$region == "SSA", "ymean"]`, whereas in other regions like East Asia and Pacific, the average year of graduation is `r avg_year[avg_year$region == "EAP", "ymean"]`. The current pace at which some countries in Africa are reducing poverty is so slow that, were they continue their current poverty-reduction trend, countries like `r cty_g[1, "countryname"]`, `r cty_g[2, "countryname"]`, and `r cty_g[3, "countryname"]` would eliminate poverty in `r cty_g[1, "year"]`, `r cty_g[2, "year"]`, and `r cty_g[3, "year"]`, respectively. 


# National poverty rate

National poverty lines are usually the monetary vaue of a basket of goods that meet certain criteria defined by countries as the essential statandard of living. Ideally, such a basket not only meets the criteria such a caloric intake, but also represents the consumption patterns of each country. Thus, the set of goods of the basket and their corresponding values are different across countries. If a household does not have enough resources to purchase the basket, it is concidered poor under national definitions. Goal 1 seeks to reduce the share of the population living in poverty by half, as defined by these national poverty lines.

Graph `x` shows the progression of poverty in countries with enough data.  The bars represent the change in poverty respect to the initial value; poverty declines if a bar rises towards the centre of the circle and increases otherwise. A country accomplishes the poverty goal reduction once the corresponding bar breaks the red circle barrier on the graph. This red circle is hence indicating a poverty reduction of half of the initial value. Within the sample, 20 countries have accomplished the goal by 2015, as shown by the darker bars. The most numerous changes take place in Europe & Central Asia with six countries, followed by South Asia and East Asia & Pacific with four nations, Latin America & Caribbean with two, and Sub-Sharan Africa and Middle East & North Africa with one.

Beyond the number of countries that have accomplished the goal, and considering the 15-year projections , indicated with the lighter colour bars, the regional performance seems to be diverse.  Despite the overall positive performance of East Asia & Pacific, some countries of this region as Micronesia, Tonga and Timor-Leste display worrying trends with an increase in poverty and, if persisting at the same course, two of them raising over half of the original value in 15 years. In other less performing regions as the Middle East & North Africa and Sub-Sharan Africa, the trends are more bothersome. In Sub-Sharan Africa, if the trend continues, only one other country may reach the goal, leaving the region trapped in poverty, while in the Middle East & North Africa no additional country seems to be close to escape poverty. Meanwhile, Latin America & Caribbean and Europe & Central Asia have more fortunate trends. Despite some countries slightly growth, most of the nations have reduced poverty meaningfully in these regions, nonetheless much the goal still far from accomplished. A similar scenario takes place in South Asia, but here Afghanistan presents an extremely worrying case. 

Long story short, there are notable cases of successful poverty reduction, being some regions more successful than others, yet, the goal still far from accomplished. Moreover, poverty reduction needs to accelerate, especially in Sub-Sharan Africa, if we intend to achieve to reduce the share of the population living in poverty by half in the foreseeable future.
