---
title: Goal 1. End poverty in all its forms everywhere
subtitle: Your subtitle
author: R.Andres Castaneda
date: Created `r Sys.Date()`
output:
  redoc::redoc:
    reference_docx: "template.docx"
    highlight_outputs: TRUE
    margins: 1 
    line_numbers: FALSE 
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  warning = FALSE,
  collapse = TRUE,
  message = FALSE,
  echo = FALSE,
  fig.retina = 3
)
knitr::opts_knit$set(
  root.dir = ".."
)
```

```{r calculations}

source("R/SDG_01_data.R")
source("R/SDG_01_charts.R")

# No. of poor
npoor <- wld %>% 
  arrange(year) %>% 
  filter(year  %in% c(1993, 2015)) %>% 
  select(poor_pop) %>% 
  round(digits = 0) %>% 
  pull()

np <- (npoor[1] - npoor[2])


# Headcounts
hc <- wld %>% 
  arrange(year) %>% 
  filter(year  %in% c(1993, 2015)) %>% 
  select(headcount) %>% 
  {round(.*100, digits = 1)} %>% 
  pull()

tmp = "TEMP"
x <- "x"

#--------- regional poveryt

r <- povcalnet_wb() 


#--------- CHN and IND poor population
np_ic <- cty2 %>% 
  filter(year  %in% c(1993, 2015)) %>% 
  group_by(year) %>% 
  summarise(pop = sum(poor_pop)) %>% 
  ungroup() %>% 
  arrange(-year) %>% 
  summarise(diff = diff(pop)) 

sh_np <- prettyNum((np_ic/10)/np , digits = 3)

```

# The story behind Global poverty

The first target of Goal 1 of the SDG is to eliminate extreme poverty by 2030.
Based on the historical trend of the most recent data, the World seems to be
going in the right direction to meet the goal. In 1990, `r hc[1]`
percent of the population (`r prettyNum(npoor[1]/1000, digits = 2)` billion people) lived with
less than \$1.9 a day (PPP 2011 prices). 25 years later,
`r prettyNum((np/npoor[1])*100, digits = 2)` percent of theme escaped extreme poverty, leaving
behind `r prettyNum(npoor[2], digits = 3)` million people (`r hc[2]` of the
population).

If we assume that the trend of poverty reduction will remain the same over the
next decade as it has been during the last 25 years, Goal 1 would be met by
**2024!**. Unfortunately, this optimistic scenario is not support by the data.
Let's see why.

First, the underlying data of the Global poverty rate in a given year are the
poverty rates of all the countries in the world. In some of them, the poverty
rate is high, but in others it is low. The dispersion of poverty rates in a
given years is so wide that scale at which global poverty is usually analized
must be increased to make room for all the poverty rates to fit the chart. At
this scale, the decreasing pace of global poverty does not seem that rapid
anymore.

How could it be then that the dispersion of poverty rates has remained
relatively constant over time and yet global poverty rate has decreased? This is
so because global poverty is a population-weigthed average of country poverty
rates. The larger the country the more it weighs in the global average.

By looking closely at the underlying data, we find that the world-wide poverty
story of the last 25 five years is mainly explained by the economic performance
of two countries: China and India. Out of the `r prettyNum(np, digits = 2)` billion
people that escaped poverty from 1990 to 2015, `r np_ic` million
(`r sh_np` percent) is due to these two countries. The other
`r 100- as.numeric(sh_np)` percent is the contribution of the rest of the world.

```{r china}
chn_mm <- cty2 %>% 
    group_by(countrycode) %>% 
    slice(which.min(year), which.max(year)) %>% 
    filter(countrycode == "CHN") %>% 
    select(headcount) %>% 
    pull()
```

China, in particular, has reduced its poverty rate from
`r prettyNum(chn_mm[[1]]*100, digits = 2)` percent in 1993 to `r prettyNum(chn_mm[[2]]*100, digits = 2)` percent
in 2015. The number of extreme poor in China is so small today that it bearly
counts in the global average. India, on the other, though it has reduced by half
its poverty rate, it still has some room to contribute to reduce global poverty.

During the same period, other countries have reduced their poverty rate as well,
but many small countries still have poverty rates that are way higher than the
world average. Unfortunately, most of those countries belong to Subsaharan
Africa and in many of them poverty has increased during during the last decade.

Other regions like Latin America and Europe and Central Asia are not doing that
bad. Most of the their countries have poverty rates below the global average and
the great majority has reduced their povety rates over the same period.

Summing up, this is how the world looks today. The higest poverty rates are
concentrated in Africa and South Asia and only two countries, India and Nigeria.
have `r x` of the poor. Yet, How do we expect it will look
like inthe future?

According to `r x`, if we assume an accelerated economic
growth and a constant reduction of inequality in all the countries, the best
projection of global poverty still fails to eliminate global poverty by 2030. In
contrast, projections of poverty rates that assume poor economic growth and
rising welfare inequality are about `r scales::percent(max(pty$headcount), suffix = "")` percent.

# Different trends across countries

If we assume that each country were to keep the same poverty-reduction trend of the last 20 years, we could divide the countries in the world in three groups. Those that will eventually eliminate poverty, those that won't eliminate poverty, and those that we jsut don't know. 

Countries whose poverty trends have been negative during the last years will eventually eliminate poverty as long as the trend remains unchanged. The current poverty-reduction trends are not predictor of poverty levels in the future, as they do not take into account poitical, economic, or exogenous events that affect poverty levels. However, the trend provides us with sense of the current pace at which each country is reducing poverty. This pace may change in the future, but if it were to continue indefinetly as is, countries that are currently reducing poverty levels will eventually eliminate poverty completely. The question then is when would that happen. 



  


# National poverty rate

