---
title: Goal 1. End poverty in all its forms everywhere
subtitle: The near future of global poverty
author: R.Andres Castaneda
date: Created `r Sys.Date()`
output:
  redoc::redoc:
    reference_docx: template.docx
    highlight_outputs: yes
    margins: 1
    line_numbers: no
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse   = TRUE,
  warning    = FALSE,
  message    = FALSE,
  echo       = FALSE,
  fig.retina = 3 
  # fig.width  = 8,
  # fig.align  = "center"
)
knitr::opts_knit$set(
  root.dir = ".."
)
```

```{r calculations}
library("english")
library("scales")

source("R/SDG_01_data.R")
source("R/SDG_01_charts.R")
source("R/utils.R")

# No. of poor
npoor <- wld %>% 
  arrange(year) %>% 
  filter(year  %in% c(1993, 2015)) %>% 
  select(poor_pop) %>% 
  round(digits = 0) %>% 
  pull()

np <- (npoor[1] - npoor[2])

# Headcounts
hc <- wld %>% 
  arrange(year) %>% 
  filter(year  %in% c(1993, 2015)) %>% 
  select(headcount) %>% 
  {round(.*100, digits = 1)} %>% 
  pull()

tmp = "TEMP"
x <- "x"

#--------- regional poverty

r <- povcalnet_wb() 

#--------- CHN and IND poor population
np_ic <- cty2 %>% 
  filter(year  %in% c(1993, 2015)) %>% 
  group_by(year) %>% 
  summarise(pop = sum(poor_pop)) %>% 
  ungroup() %>% 
  arrange(-year) %>% 
  summarise(diff = diff(pop)) %>% 
  pull()

sh_np <- percent(np_ic/np, suffix = "", accuracy = 0.1)

```

The story behind Global poverty
===============================

The first target of Goal 1 to eliminate extreme poverty by 2030. Based on the
historical trend of the most recent data, the World seems to be going in the
right direction to meet the goal. In 1990, `r hc[1]` percent of
the population (`r prettyNum(npoor[1]/1000, digits = 2)` billion people) lived with less than
\$1.9 a day (2011 PPP prices). 25 years later, `r prettyNum((np/npoor[1])*100, digits = 2)` percent
of them escaped extreme poverty, leaving behind `r prettyNum(npoor[2], digits = 3)` million
people (`r hc[2]` of the population).

If we assume that the trend of poverty reduction will remain the same over the
next decade as it has been during the last 25 years, Goal 1 would be met by
**2024!**. Unfortunately, this optimistic scenario is not support by the data.
Let's see why.

First, the underlying data of the Global poverty rate in a given year are the
poverty rates of all the countries in the world. In some of them, poverty rates
whereas in others they are low. The dispersion of country-poverty rates in a
given year is so wide that the scale at which global poverty is usually analyzed
must be enlarged to make room for all the poverty rates to fit in the chart. At
this scale, the decreasing pace of global poverty does not seem that rapid
anymore.

How could it be then that the dispersion of poverty rates has remained
relatively constant over time and yet global poverty rate has decreased? This is
so because global poverty is a population-weighted average of country poverty
rates. The larger the country the more it weighs in the global average.

By looking closely at the underlying data, we find that the global poverty story
of the last 25 five years has been driven mainly by the economic performance of
two countries: China and India. Out of the `r prettyNum(np/1000, digits = 3)` billion
people that escaped poverty from 1990 to 2015, `r np_ic` million
(`r sh_np` percent) is due to these two countries. The other
`r 100- as.numeric(sh_np)` percent is the contribution of the rest of the world.

```{r china}
chn_mm <- cty2 %>% 
    group_by(countrycode) %>% 
    slice(which.min(year), which.max(year)) %>% 
    filter(countrycode == "CHN") %>% 
    select(headcount) %>% 
    pull()
```

China, in particular, has reduced its poverty rate from
`r prettyNum(chn_mm[[1]]*100, digits = 2)` percent in 1993 to `r prettyNum(chn_mm[[2]]*100, digits = 2)` percent
in 2015. The number of extreme poor in China is so small today that it barely
counts in the global average. India, on the other, though it has reduced by half
its poverty rate, it still has some room to contribute to reduce global poverty.

During the same period, other countries have reduced their poverty rate as well,
but many small countries still have poverty rates that are way higher than the
world average. Unfortunately, most of those countries belong to Sub-Saharan
Africa and in many of them poverty has increased during the last decade.

Other regions like Latin America and Europe and Central Asia are not doing that
bad. Most of the their countries have poverty rates below the global average and
the great majority has reduced their poverty rates over the same period.

This is how the world looks today. The highest poverty rates are concentrated in
Africa and South Asia and only two countries, India and Nigeria, have
`r x` of the poor today. Then, how do we expect it will look
like in the future?

According to `r x`, if we assume an accelerated economic
growth and a constant reduction of inequality in all the countries, the best
projection of global poverty still fails to eliminate global poverty by 2030. In
contrast, projections of poverty rates that assume poor economic growth and
rising welfare inequality are about `r percent(max(pty$headcount), suffix = "")` percent.

Different trends across countries
=================================

If we assume that each country was to keep the same poverty-reduction trend of
the last 20 years, we could divide the countries in the world in three groups.
Those that will eventually eliminate poverty, those that won't eliminate
poverty, and those that we just don't know.

Countries whose poverty trends have been negative during the last years will
eventually eliminate poverty as long as the trends remain unchanged. The current
poverty-reduction trends are not a predictor of poverty levels in the future, as
they do not take into account poitical, economic, or exogenous events that
affect poverty levels, but the trend provides us with a sense of the current
pace at which each country is reducing poverty. This pace may change in the
future, but if it were to continue indefinitely as is, countries that are
currently reducing poverty levels will eventually eliminate poverty. The
question is then, when would that happen?

```{r graduation}
lower_year <- 2020

cty_g <- cty_p %>% 
  filter(year > lower_year) %>% 
  arrange(-year) 

shr_grd <- tabyl(cty_g, region)  # graduation share

grd_per <- shr_grd %>%           # Africa share
  mutate(
    percent = percent(percent, 
                      suffix = "", 
                      accuracy = 0.1)
    ) %>% 
  select(-n)

avg_year <- cty_g %>%        # average year
  group_by(region) %>% 
  summarise(ymean = floor(mean(year)))

worst_cty <- cty_g %>% 
  slice(1:3) %>% 
  select(countryname, year) %>% 
  mutate(
    year = as.character(year)
  )

```

Among all the countries that eventually would eliminate poverty after
`r lower_year`, `r grd_per[grd_per$region == "SSA", "percent"]` percent are from Sub
Saharan Africa, `r grd_per[grd_per$region == "LAC", "percent"]` percent are from Latin America, and
`r grd_per[grd_per$region == "EAP", "percent"]` percent are from East Asia and Pacific. This
proportion is expected as the number of countries in Africa is way larger than
in any other region, but what is unexpected is their year of graduation. In
average, African countries would graduate in `r avg_year[avg_year$region == "SSA", "ymean"]`, whereas
in other regions like East Asia and Pacific, the average year of graduation is
`r avg_year[avg_year$region == "EAP", "ymean"]`. At least in these two countries, the average country
won't meet the goal by 2030. The current pace at which some countries in Africa
are reducing poverty is so slow that, were they to continue their current
poverty-reduction trend, countries like `r add_and(worst_cty[["countryname"]])` would
eliminate poverty in `r add_and(worst_cty[["year"]])`, respectively. Figure
`r x` shows the graduation year of a subset of countries.

```{r badcountries}

wrs_reg <- bad_ctrs %>% 
  tabyl(regionname) %>% 
  filter(n == max(n)) %>% 
  select(regionname) %>% 
  pull


wrs_regn <- bad_ctrs %>% 
  tabyl(regionname) %>% 
  summarise(
    nsum = sum(n),
    nmax = max(n)
  )


wrs3 <- bad_ctrs %>% 
  arrange(-headcount)  %>% 
  slice(2:4) %>% 
  select(countryname) %>% 
  pull() %>% 
  add_and()

wrs_gp <- bad_ctrs %>% 
  arrange(-gr_pp) %>% 
  slice(1:2) %>% 
  transmute(
    countryname = countryname, 
    gr_pp       = percent(gr_pp, suffix = "", accuracy = 0.1)
    )
```


though the graduation year for some of the countries above is more than `r paste(as.english(150))` years from now, some other countries won't ever graduate if they continue with the same poverty trend of the last two decades. `r wrs_reg` has largest share of these countries wirh `r wrs_regn[["nmax"]]` our of `r wrs_regn[["nsum"]]`. Among them, we find some of the poorest economies of the world like `r wrs3` and, even worst off,  countries like `r add_and(wrs_gp[["countryname"]])` have the highest annualized growth of the number of poor, with `r add_and(wrs_gp[["gr_pp"]])` percent, respectively . 

National poverty rate
=====================

```{r natline}
source("R/SDG_01_NatLines.R")
```

National poverty lines are usually the monetary value of a basket of goods that
meet certain criteria---usually caloric intake---defined by countries as the
essential standard of living. Ideally, such a basket not only meets the criteria
but is also representative of the consumption patterns of a society, making the
set of goods of the basket and their corresponding values different across
countries. If a household does not have enough resources to purchase the basket,
it is considered poor under national definitions. Goal 1 seeks to reduce the
share of the population living in poverty by half, as defined by these national
poverty lines.

```{r natcalc}
nat_dec <- tabyl(overall, dec) %>%           # Africa share
  mutate(
    percent = percent(percent, 
                      suffix = "", 
                      accuracy = 0.1)
  )

n_goal <- overall %>% 
  filter(Growthp < -.5) %>%        # met the goal
  summarise(n_distinct(countrycode)) %>% 
  pull()

reg_goal <- overall %>% 
  filter(Growthp < -.5) %>%        
  tabyl(regioncode)

# failing countries 
failing <- overall %>% 
  filter(Growthp > 0) %>% 
  arrange(regioncode, -Growthp) %>% 
  select(regioncode, countrycode, countryname, Growthp)

eap_cty <- failing %>% 
  filter(regioncode == "EAP") %>% 
  select(countryname) %>% 
  pull() %>% 
  add_and()

n_eap <- paste(as.english(reg_goal[reg_goal$regioncode == "ECA", "n"]))
n_sar <- paste(as.english(reg_goal[reg_goal$regioncode == "SAS", "n"]))

```

Figure `r x` shows the progression of poverty in countries
with enough data. The bars represent the change in poverty respect to the
initial value. Poverty declines if a bar rises towards the center of the circle
and increases otherwise. A country accomplishes the poverty goal reduction once
the corresponding bar cross the red circle barrier on the graph, which indicates
a poverty reduction of half of the initial value. Within the sample,
`r n_goal` countries have accomplished the goal by
`r year2`, as shown by the darker bars. The most numerous
changes take place in Europe & Central Asia with `r n_eap`
countries, followed by South Asia with `r n_sar`.

Beyond the number of countries that have accomplished the goal, and considering
the `r yspam`-year projections, indicated with the lighter color
bars, the regional performance seems to be diverse. Despite the overall positive
performance of East Asia & Pacific, some countries of this region as
`r eap_cty` display worrying trends with an increase in poverty
and two of them raising over half of the original value in
`r yspam` years from now. In other less performing regions like
the Middle East and North Africa and Sub-Sharan Africa, the trends are more
bothersome. In Sub-Sharan Africa, if the trend continues, only one of two
countries may reach the goal, leaving the region trapped in poverty. In the
Middle East & North Africa no additional country seems to be close to escape
poverty. Meanwhile, Latin America & Caribbean and Europe & Central Asia have
more fortunate trends. Despite the modest growth of some countries, most of the
nations have reduced poverty meaningfully. Nonetheless, we are still far from
accomplishing the goal. A similar scenario takes place in South Asia, but here
Afghanistan presents an extremely worrying case.

In short, there are notable cases of successful poverty reduction, in which some
regions are more successful than others. Yet, the goal is still far from being
accomplished. Poverty reduction needs to accelerate, especially in Sub-Sharan
Africa, if we intend to reduce the share of the population living in poverty by
half in the foreseeable future.
