---
title: Goal 10. Reduce inequalities
subtitle: 
author: R.Andres Castaneda
date: Created `r Sys.Date()`
output:
  redoc::redoc:
    reference_docx: template.docx
    highlight_outputs: yes
    margins: 1
    line_numbers: no
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse   = TRUE,
  warning    = FALSE,
  message    = FALSE,
  echo       = FALSE,
  fig.retina = 3 
  # fig.width  = 8,
  # fig.align  = "center"
)
knitr::opts_knit$set(
  root.dir = ".."
)
```

```{r calculations}
library("english")
library("scales")
library("tidyverse")
library("ggplot2")
library("ggpubr")

source("R/SDG_10_Growth40_2017rep.R")

WDI <- WDI %>% 
  mutate(Ddiff = diff - OLDdiff)

Group <- c("regionname", "incomegroup")

for (i in Group){
  groupd <- WDI %>% 
    mutate(gr = if_else(diff >= 0,1,0),
           ls = if_else(diff < 0,1,0),
           pos = if_else(Growth >= 0,1,0),
           neg = if_else(Growth < 0,1,0),
           pos40 = if_else(Growth40 >= 0,1,0),
           neg40 = if_else(Growth40 < 0,1,0),
           grOLD = if_else(OLDdiff >= 0,1,0),
           lsOLD = if_else(OLDdiff < 0,1,0),
           posOLD = if_else(OLDGrowth >= 0,1,0),
           negOLD = if_else(OLDGrowth < 0,1,0),
           pos40OLD = if_else(OLDGrowth40 >= 0,1,0),
           neg40OLD = if_else(OLDGrowth40 < 0,1,0),
           counter = if_else(!is.na(diff), 1,0) ,
           counterOLD = if_else(!is.na(OLDdiff), 1,0) 
    )%>% 
    group_by_(i) %>% 
    summarise(gr = sum(gr, na.rm = T),
              ls = sum(ls, na.rm = T),
              pos = sum(pos, na.rm = T),
              neg = sum(neg, na.rm = T),
              pos40 = sum(pos40, na.rm = T),
              neg40 = sum(neg40, na.rm = T),
              grOLD = sum(grOLD, na.rm = T),
              lsOLD = sum(lsOLD, na.rm = T),
              posOLD = sum(posOLD, na.rm = T),
              negOLD = sum(negOLD, na.rm = T),
              pos40OLD = sum(pos40OLD, na.rm = T),
              neg40OLD = sum(neg40OLD, na.rm = T),
              cc = sum(counter, na.rm = T),
              ccOLD = sum(counterOLD, na.rm = T)
    ) %>%
    mutate(grr = round(gr / cc *100, digits = 2),
           lsr = round(ls/cc * 100, digits = 2),
           grrOLD = round(grOLD / ccOLD * 100, digits = 2),
           lsrOLD = round(lsOLD/ccOLD*100, digits = 2)
           ) 
  
  assign(i, groupd)
  rm(groupd)
  
  # rank 
  
}

options(scipen=1, digits=2)

```

The bottom 40
===============================

Goal 10 aims to lessen inequalities within the citizens of a nation, promoting wealth and sustained wellbeing for everyone. On that line, the fastest-growing incomes should be the ones of the most vulnerable. In `r nrow(WDI[WDI$diff > 0 & !is.na(WDI$diff),])` out of `r nrow(WDI[!is.na(WDI$diff),])` countries, with enough data, this is true, in these nations the earnings of poorest 40 per cent of the population grew over the national average.

Besides the considerable improvement in several countries, inequalities persist in many others. The income of the bottom 40 increase was less than the national average in `r nrow(WDI[WDI$diff < 0 & !is.na(WDI$diff),])` countries, of which in `r nrow(WDI[WDI$diff < 0 & WDI$Growth40 < 0 & WDI$Growth > 0 & !is.na(WDI$diff),])` their income decreased, while the average did not. Interestingly, out of the `r nrow(WDI[WDI$Growth < 0 & !is.na(WDI$diff),])` countries with an average decrease in incomes, in `r nrow(WDI[WDI$diff < 0 & WDI$Growth < 0 & !is.na(WDI$diff),])` the bottom 40 were more harshly hit, meanwhile in `r nrow(WDI[WDI$Growth40 >= 0 & WDI$Growth < 0 & !is.na(WDI$diff),])` they achieved a welfare increase. 

In a regional perspective, the bottom 40 of most of the countries of `r paste(str_c(arrange(regionname[regionname$grr>=50,], desc(grr))$regionname[1:nrow(regionname[regionname$grr>=50,])-1], collapse = ", "), arrange(regionname[regionname$grr>=50,], desc(grr))$regionname[nrow(regionname[regionname$grr>=50,])], sep =  ", and ")` increased their income over the average; meanwhile, in `r paste(str_c(arrange(regionname[regionname$grr<50,], desc(lsr))$regionname[1:nrow(regionname[regionname$grr<50,])-1], collapse = ", "), arrange(regionname[regionname$grr<50,], desc(lsr))$regionname[nrow(regionname[regionname$grr<50,])], sep =  ", and ")`  their income improved less than the national average, with `r paste0(str_c(arrange(regionname[regionname$grr<50,], desc(lsr))$lsr[1:nrow(regionname[regionname$grr<50,])-1], collapse = "%, "), "% and ", paste0(arrange(regionname[regionname$grr<50,], desc(lsr))$lsr[nrow(regionname[regionname$grr<50,])],"%"))` of their countries exhibiting such behaiviour, respectively.

An overview by income groups is largely meaningful. The `r paste0(str_c(arrange(incomegroup[incomegroup$grr>=50,], desc(lsr))$grr[1:nrow(incomegroup[incomegroup$grr>=50,])-1], collapse = "%, "), "% and ", paste0(arrange(incomegroup[incomegroup$grr>=50,], desc(lsr))$grr[nrow(incomegroup[incomegroup$grr>=50,])],"%"))` of  `r paste(str_c(arrange(incomegroup[incomegroup$grr>=50,], desc(grr))$incomegroup[1:nrow(incomegroup[incomegroup$grr>=50,])-1], collapse = ", "), arrange(incomegroup[incomegroup$grr>=50,], desc(grr))$incomegroup[nrow(incomegroup[incomegroup$grr>=50,])], sep =  ", and ")`'s bottom 40% increased their income more rapidly than de average, accordingly; Whereas, for the `r paste(str_c(arrange(incomegroup[incomegroup$grr<50,], desc(lsr))$incomegroup[1:nrow(incomegroup[incomegroup$grr<50,])-1], collapse = ", "), arrange(incomegroup[incomegroup$grr<50,], desc(lsr))$incomegroup[nrow(incomegroup[incomegroup$grr<50,])], sep =  ", and ")` countries the bottom earners experienced a lower expansion of their welfare compared to the average, with only `r paste0(str_c(arrange(incomegroup[incomegroup$grr<50,], desc(lsr))$grr[1:nrow(incomegroup[incomegroup$grr<50,])-1], collapse = "%, "), "% and ", paste0(arrange(incomegroup[incomegroup$grr<50,], desc(lsr))$grr[nrow(incomegroup[incomegroup$grr<50,])],"%"))` of the countries, in the respective income group, with the bottom's growth out-performing the average. The full information on the countries with enough data is summarized on the panel A of figure 1; the solid red dots represents the bottom's growth, meanwhile the hollow black circle, represents the averages. 

```{r, fig.width = 12}
# Small fig.width
p1s <- p1s + ggtitle("") + labs(caption = paste(""))
p2s <- p2s + ggtitle("") + labs(caption = paste(""))
p3s <- p3s + ggtitle("")
main <- ggarrange(p1s, p2s, p3s,
                    ncol = 3, nrow = 1)  %>% 
  annotate_figure(top = text_grob("Growth Bottom 40 vs Average", face = "bold", size = 14))
plot(main)
```

It can be valuable to check how the countries have performed across time. Comparing the current growth rates (2012-2017) with the ones between 2008-2013, we found broad changes in the countries' behaviour. Panel B of the figure 1, shows the growth rates for the countries in which for the 2008-2013 period their bottom grew faster than the average; here the light blue figures correspond to the values for the 2008-2013 period. One of the first conclusions to arise, is that for most countries the behaviour is volatile across time. Out of the `r nrow(WDI[WDI$OLDdiff>0 & !is.na(WDI$diff & WDI$OLDdiff),])` countries were that previously experiencing a fastest growth for the bottom, only `r nrow(WDI[WDI$OLDdiff>0 & WDI$diff>0 & !is.na(WDI$diff & WDI$OLDdiff),])` endured. Moreover, the changes in a countries' overall growth seems to drive inequality changes. Out of the `r nrow(WDI[WDI$OLDdiff>0 & WDI$OLDGrowth<0 & WDI$Growth>0 & !is.na(WDI$diff & WDI$OLDdiff),])` countries with faster growth for the bottom in the past and experiencing an average recovery from a previous negative growth, `r nrow(WDI[WDI$OLDdiff>0 & WDI$OLDGrowth<0 & WDI$Growth>0 & WDI$diff < 0 & !is.na(WDI$diff & WDI$OLDdiff),])` reverted their former inequality reducing trends. A similar shift is found when checking for an overall increase in the average income; in `r nrow(WDI[WDI$OLDdiff>0 & WDI$OLDGrowth<WDI$Growth & WDI$diff < 0 & !is.na(WDI$diff & WDI$OLDdiff),])` countries, out of `r nrow(WDI[WDI$OLDdiff>0 & WDI$OLDGrowth<WDI$Growth & !is.na(WDI$diff & WDI$OLDdiff),])` perciving an increase in their mean growth, the trend changed, and the bottom 40 started to displays slower growth compared to the mean. While, in the `r nrow(WDI[WDI$OLDdiff>0 & WDI$OLDGrowth>WDI$Growth & !is.na(WDI$diff & WDI$OLDdiff),])` countries with a reduction on the mean growth, in only `r nrow(WDI[WDI$OLDdiff>0 & WDI$OLDGrowth>WDI$Growth & WDI$diff < 0 & !is.na(WDI$diff & WDI$OLDdiff),])` the bottom's growth became slower.

Finally, for both growth rate windows, the following figure shows how the countries rank from the one with the fastest growth of the bottom compared to the average, to the one with the slowest. The black points represent the rank during the 2008-2013 period, while the colored points are the new rankings. The color of the new points depends on whether the country goes up in the rank (blue), or down (red). Once again, the temporal behaviour shows volatile, with no country remaining on the same position. The countries with the smallest relative change were `r str_c(head(arrange(rankd1, abs(Drank))$country, n = 2), collapse = " and ")`, and those with the largest were `r str_c(tail(arrange(rankd1, abs(Drank))$country, n = 2), collapse = " and ")`, ranking `r str_c(tail(arrange(rankd1[rankd1$Drank < 0,], abs(Drank))$rankNEW, n = 2), collapse = " and ")` for the latest data, compared to a previous rank of `r str_c(tail(arrange(rankd1[rankd1$Drank < 0,], abs(Drank))$rankOLD, n = 2), collapse = " and ")`, respectivly. For `r nrow(rankd1[rankd1$Drank > 0 & rankd1$diffstatus == "Faster",])` countries, of `r nrow(rankd1[rankd1$Drank > 0,])`  going up in the rank, the bottom's growth compared to the average improved; even if in `r nrow(rankd1[rankd1$Drank > 0 & rankd1$diffstatus == "Faster" & rankd1$diff < 0,])` cases their growth still slower than the mean. At the same time, for the countries that worsen their ranking, `r ifelse(nrow(rankd1[rankd1$Drank < 0 & rankd1$diffstatus == "Faster",])==0, " no country", paste0(nrow(rankd1[rankd1$Drank < 0 & rankd1$diffstatus == "Faster",])," countries"))` improved their bottom's growth.



```{r, fig.width = 14}
plot(plot_rank_s1)
```

