---
title: Goal 10. Reduce inequalities
subtitle: 
author: R.Andres Castaneda
date: Created `r Sys.Date()`
output:
  redoc::redoc:
    reference_docx: template.docx
    highlight_outputs: yes
    margins: 1
    line_numbers: no
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse   = TRUE,
  warning    = FALSE,
  message    = FALSE,
  echo       = FALSE,
  fig.retina = 3 
  # fig.width  = 8,
  # fig.align  = "center"
)
knitr::opts_knit$set(
  root.dir = ".."
)
```

```{r calculations}
library("english")
library("scales")
library("tidyverse")

source("R/SDG_10_Growth40.R")

# of countries w/ 40 > 100 
wdata <- WDI %>% 
  mutate(Dgrowth = Growth40 - Growth) %>% 
  arrange(Dgrowth)

gr <- wdata %>% 
  filter(Dgrowth >= 0)

ls <-wdata %>% 
  filter(Dgrowth < 0)

ls_pos <- ls %>% 
  filter(Growth40>=0)

ls_neg <- ls %>% 
  filter(Growth40<0)

ls_negGP <- ls_neg %>% 
  filter(Growth40<0 & Growth>=0)

absneg <- wdata %>% 
  filter(Growth40 < 0)

pos <- wdata %>% 
  filter(Growth >= 0)

neg <- wdata %>% 
  filter(Growth < 0)


regions <- wdata %>% 
  mutate(gr = if_else(Dgrowth >= 0,1,0)) %>%
  mutate(ls = if_else(Dgrowth < 0,1,0)) %>%
  mutate(pos = if_else(Growth >= 0,1,0)) %>%
  mutate(neg = if_else(Growth < 0,1,0)) %>%
  mutate(pos40 = if_else(Growth40 >= 0,1,0)) %>%
  mutate(neg40 = if_else(Growth40 < 0,1,0)) %>%
  mutate(counter = 1) %>% 
  group_by(region) %>% 
  summarise(gr = sum(gr), ls = sum(ls), pos = sum(pos), neg = sum(neg), pos40 = sum(pos40), neg40 = sum(neg40), cc = sum(counter)) %>% 
  mutate(grr = gr / cc, lsr = ls/cc)


```

The bottom 40
===============================

Goal 10 aims to lessen inequalities within the citizens of a nation, promoting wealth and sustained wellbeing for everyone. On that line, the fastest-growing incomes should be the ones of the most vulnerable. In `r nrow(gr)` out of `r nrow(wdata) ` countries, with enough data, this is true, in these nations the earnings of poorest 40 per cent of the population grew over the national average.

Besides the considerable improvement in several countries, inequalities persist in many others. The income of the bottom 40 increase was less than the national average in `r nrow(ls)` countries, of which in `r nrow(ls_negGP)` their income decreased, while the average did not. Interestingly, out of the `r nrow(neg)` countries with an average decrease in incomes, in `r sum(neg$Dgrowth < 0)` the bottom 40 were more harshly hitten, meanwhile in `r sum(neg$Growth40 >= 0)` they achieved a welfare increase. 

In a regional perspective, the bottom 40 of most of the countries of EAP, ECA and LAC increased their income over the average; meanwhile, in OHI, SSA, MNA and SAS  their income improved less than the national average.  The SAS case is worrying, without a single country where the bottom's income grew more than the average. Also, OHI behaviour is troubling as inequality seems to be increasing despite the overall increase in incomes. 

