---
title: "SDG 10"
output:
  xaringan::moon_reader:
    seal: false
    transition: fade
    css: ["css/sfah.css", "css/fonts.css", "default"]
    self_contained: false
    lib_dir: libs
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      titleSlideClass: ["left", "middle", "inverse"]
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(knitr)
library(scales)
library(ggpubr)
library(here)

knitr::opts_chunk$set(
  warning = FALSE,
  collapse = TRUE,
  message = FALSE,
  echo = FALSE,
  fig.retina = 3,
  fig.width = 10
)
knitr::opts_knit$set(
  root.dir = ".."
)
# options(knitr.table.format = "html")
library(fontawesome) # from github: https://github.com/rstudio/fontawesome devtools::install_github("rstudio/fontawesome")
xaringanExtra::use_xaringan_extra(c("tile_view", "animate_css", "tachyons")) # devtools::install_github("gadenbuie/xaringanExtra")
```

class: inverse, right, bottom
## SDG Atlas 2020

## Goal 10: Reduce inequality within and among countries

???
  Intcomments

---
class: inverse, right, bottom
### Non-monetary inequality
```{r distsim}
if (length(ls(pattern = "ld1")) == 0 ) {
  source(here("R", "get_WDI.R"))
  source(here("R", "SDG_10_non_monetary_Growth.R"))
  ld1 <-  1
}

```
---
class: middle, center

### Social Protection Policies
---
.left-column[
For 2007 social safety net progams had the following coverage per income quintiles.
]
.right-column[

```{r, fig.width = 15, fig.asp=0.3}
plot(SPcover2)
```
]
---
.left-column[
In a lapse of ten years, the coverage increased in most of the countries, and segments. 

The poorest are not attaining coverage much faster than the wealthy. For the poorest, the average annual coverage growth was 3.06%, meanwhile for the richest was 2.59%. 
]
.right-column[
```{r, fig.width = 15, fig.asp=0.3}
plot(SPcover)
```

```{r, fig.width = 15, fig.asp=0.3}
plot(SPcover3)
```

]
---
.left-column[
Coverage aside, countries seem to be improving the redistributive scope of the social safety net programs.

With a few exceptions, nations that were allocating less of the social safety net benefits to the poorest 20% of the population favoured the fastest redirection of the resources towards this segment of the population.

Despite the overall tendency, countries as Côte d'Ivoire and Nigeria raise attention with low incidence to the poor and without signs of improvement.
]
.right-column[
```{r, fig.width = 12, fig.asp=0.6}
plot(SPinc)
```
]
---
.pull-left[
Despite the advancement, there is plenty of room for social protection policies

The Country Policy and Institutional Assessment (CPIA) intends to capture the quality of a country’s policies and institutional arrangements, focusing on central elements under the country’s control, and rates the countries between 1 (low) and 6 (high).

The development of assessments on social protection shows a great improvement in countries as Zimbabwe, Vanuatu and Togo. However, in a lapse of ten years, the rating is unchanged for many counties, with some of the most vulnerable territories as Haiti and Congo experiencing a rating downgrade. 
]
.pull-right[
```{r, fig.width = 7, fig.asp=1.2}
plot(SPcpi)
```
]
---
class: middle, center

### Remittances
---
.left-column[
During 2017, an estimate of 585.93 billion dollars was sent as remittances across the globe. From the total of remittances 51.55% were sent to low income and lower-middle-income countries; being the lower-income countries the top recipients.

Furthermore, the cost of remittances is high. For 2017 the cost of sending 200 USD was on average 7.96% of the sent value. 

Workers abroad from low-income countries as Angola, Botswana and Mozambique are the ones facing the higher cost to send money back to their families.
]
.right-column[
```{r, fig.width = 12, fig.asp=0.7}
plot(premit)
```
]
---
.left-column[
Even if high, the remittance cost in 2017 is an improvement from the 2011's average cost of 9.66%. 

Between 2011 and 2017 the remittance cost was reduced by an average of 3.93% of its value every year. 

Most of the countries experienced a cost reduction, considering both origin and destinations. However, the cost of the remittances from some of the wealthiest Arab countries with high amounts of foreign workers as the United Arab Emirates, Qatar, and Saudi Arabia is increasing. ]
.right-column[
```{r, fig.width = 12, fig.asp=0.7}
plot(premitc)
```
]
---
class: middle, center

### Social Inclusion
---
.pull-left[
.grey[Social inclusion and non-discrimination by gender or race is an essential element of an equitable society. 

The CPIA assessment of social inclusion rates countries policies from 1 (low) to 6 (high) depending on the quality of the policies implemented. For the countries with such rating, the most show moderate advancement in their equity policies, but there seem to be plenty of space for improvement.

Countries as Yemen, Madagascar and Guyana raise concerns.]
]
.pull-right[
```{r, fig.width = 7, fig.asp=1.2}
plot(pincCPI)
```
]
---
class: middle, center

### International Aid
---
.left-column[
International assistance and aid are vital to improving equity across nations. 

Between 2007 and 2017, the average net received aid increased by 3.18% annually. 

Low-income countries observed the most rapid increase with a mean annual increase of 7.03%, receiving an average of 1.46 billion in 2017.
]
.right-column[
```{r, fig.width = 12, fig.asp=0.75}
plot(pAID)
```
]
---

